---
title: ''
output: md_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 8)
```

### Estimate boxer win rates using Empirical Bayes

```{r}
library(tidyverse)
# devtools::install_github("dgrtwo/ebbr")
library(ebbr)
```


### Load scraped boxer ratings from boxrec.com

The code to scrape the boxer ratings with wins, loses, and draws is in boxrec_scraping.Rmd

```{r}
boxing <- read_csv("data/boxing_ratings.csv")

boxing <- boxing %>%
    mutate(name = str_remove_all(name, "\\*"),
           name = str_squish(name))
```

<br />

### How are number of fights, wins, loses, draws, and win rate distributed?

```{r}
boxing %>%
    select(n, wins, loses, draws, win_rate) %>%
    pivot_longer(cols = everything()) %>% 
    drop_na() %>%
    ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~name, scales = "free") +
    labs(x = NULL,
         y = "Count") +
    scale_y_continuous(labels = scales::comma_format())
```

<br />

### Do orthodox boxers have a higher win rate than southpaw boxers?

```{r}
boxing %>%
    drop_na() %>%
    ggplot(aes(stance, win_rate)) +
    geom_boxplot() +
    labs(x = "Stance",
         y = "Win rate",
         title = "The win rate between orthodox and southpaw boxers are very comprable")
```

<br />

### Who are the best boxers (i.e., those with the highest win rates)?

There are boxers who have won 100% of their boxing matches, yet they may have only boxed a small number of matches. I'll use empirical Bayesian techniques to better estimate the win rates of these 15k+ boxers.

```{r}
prior <- boxing %>%
    drop_na(wins, n) %>%
    filter(n > 10) %>%
    ebb_fit_prior(wins, n)

prior

boxing_augment <- boxing %>%
    drop_na(wins, n) %>%
    filter(n > 2) %>%
    add_ebb_estimate(wins, n, prior_subset = n > 10)

boxing_augment %>%
    ggplot(aes(win_rate, .fitted, size = n)) +
    geom_point() +
    geom_hline(yintercept = tidy(prior)$mean, color = "red", lty = 2) +
    labs(x = "Win rate",
         y = "Empirical Bayes fit",
         size = "Number of boxing matches")

```

The plot above shows how each boxer's win rate plotted against the empirical Bayes estimate. The size of each point represents how many boxing matches the boxer has had. The Bayesian estimate for boxer's with very low or very high win rates are shifted towards the average win rate (the dotted red line), unless the number of boxing matches they have had is large enough. The most drastic differences between the raw win rate and the empirical Bayesian approach is for those boxers who have very low win rates.

<br / >

```{r}
library(glue)

top_n <- 35

boxing_augment %>%
    slice_max(.fitted, n = top_n) %>%
    mutate(name = fct_reorder(name, .fitted)) %>%
    ggplot(aes(.fitted, name)) +
    geom_point() +
    geom_point(aes(x = win_rate), color = "red") +
    geom_errorbar(aes(xmin = .low, xmax = .high), width = 0.25) +
    labs(x = "Empirical bayes fit",
         y = NULL,
         title = glue("Empirical Bayes estimate of win rate for top {top_n} boxers"),
         subtitle = "Red points show win rate, error bars show 95% credible region")

```

The plot above shows the empirical Bayes estimate along with the 95% credible region for the top 35 boxers. The red dots show the boxer's win rate. The plot shows that most of the top boxer's based on the empirical Bayes estimate are currently undefeated. However, this analysis shows a true win rate of 100% is unlikely. Note: this analysis is only for active professional boxers, hence, no Floyd Mayweather who has a 100% win rate over 50 matches.

The plot below shows the same idea but for the 35 boxers with the lowest win rate estimate.

```{r}
boxing_augment %>%
    slice_min(.fitted, n = top_n) %>%
    mutate(name = fct_reorder(name, .fitted)) %>%
    ggplot(aes(.fitted, name)) +
    geom_point() +
    geom_point(aes(x = win_rate), color = "red") +
    geom_errorbar(aes(xmin = .low, xmax = .high), width = 0.25) +
    labs(x = "Empirical Bayes fit",
         y = NULL,
         title = glue("Empirical Bayes estimate of win rate for bottom {top_n} boxers"),
         subtitle = "Red points show win rate, error bars show 95% credible region")

```

<br / >

### Which boxers have estimated win rates gerater than 90%?

To determine which boxers win rate estimate are greater than 90%, I will compute the posterior error probability for each boxer. Using an alpha of 0.05, I can be 95% confident that the boxers included below have a true win rate of 90%. 

```{r}
test_prop <- 0.90

boxing_augment %>%
    add_ebb_prop_test(test_prop, alternative = "greater", sort = TRUE) %>%
    filter(.qvalue < 0.05) %>%
    mutate(name = fct_reorder(name, .fitted)) %>%
    ggplot(aes(.fitted, name)) +
    geom_point() +
    geom_errorbar(aes(xmin = .low, xmax = .high)) +
    geom_vline(xintercept = test_prop, color = "red", lty = 2) +
    labs(x = "Empirical Bayes estimate",
         y = NULL)

```

Out of `r nrow(boxing) %>% scales::comma()` boxers in the dataset, only `r boxing_augment %>% add_ebb_prop_test(test_prop, alternative = "greater", sort = TRUE) %>% filter(.qvalue < 0.05) %>% nrow()` boxers have an estimated win rate of 90+%, compared to the `r boxing %>% filter(win_rate > 0.9) %>% nrow() %>% scales::comma()` boxers that have raw win rates of 90+%. By using empirical Bayes we can get a less biased and more informative estimate of boxer's win rates (especially those who have had few matches).

<br />
<br />
<br />
<br />
<br />
<br />